---
title: "Practical Exam 1"
author: "Barbie Joy Obas"
date: "2024-03-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#A. Load the built-in warpbreaks dataset.
data(warpbreaks)

############
#A.1. Find out, in a single command, which columns of warpbreaks are either numeric or integer. What are the data types of each column?
```{r}
numeric_col <- sapply(warpbreaks, is.numeric)
numeric_col

#The data type of each column are breaks, wool and tension.
```

###########
#A.2. How many observations does it have?
```{r}
observations <- nrow(warpbreaks)
observations

#There are 54 observations
```
###########
#A.3. Is numeric a natural data type for the columns which are stored as such? Convert to integer when necessary.
```{r}

integer_col <- sapply(warpbreaks, is.integer) 
integer_col

numeric_or_integer_columns <- warpbreaks[, numeric_columns | integer_columns] 
numeric_or_integer_columns
```
###########
#A.4. Error messages in R sometimes report the underlying type of an object rather than the user-level class. Derive from the following code and error message what the underlying type. Explain what is the error all about. Do not just copy the error message that was displayed.
```{r}
#The error is an object 'numeric_or_integer_columns' not found suggests that R couldn't find the object named numeric_or_integer_columns. This error commonly occurs when you try to reference an object that hasn't been defined or is out of scope.
```
##################################################################
#B. Load the exampleFile.txt

```{r}
data <- read.csv("exampleFile.txt")
```
###########
#B.1. Read the complete file using readLines.
```{r}
lines <- readLines("exampleFile.txt")

cat(lines, sep = "\n")
```
###########
#B.2. Separate the vector of lines into a vector containing comments and a vector containing the data. Hint: use grepl.

```{r}
comments <- lines[grepl("^//", lines)]
comments

data_lines <- lines[!grepl("^//", lines)]
data_lines
```
###########
#B.3. Extract the date from the first comment line and display on the screen “It was created data.”
```{r}
date <- gsub("^// Survey data. Created : ", "", comments[1])
date
```
######################
#B.4. Read the data into a matrix as follows.

#4.a. Split the character vectors in the vector containing data lines by semicolon (;) using strsplit.
```{r}
split_data <- strsplit(data_lines, ";")
split_data
```
###########
#4.b. Find the maximum number of fields retrieved by split. Append rows that are shorter with NA's.
```{r}
max_fields <- max(sapply(split_data, length))
max_fields

split_data <- lapply(split_data, function(x) c(x, rep(NA, max_fields - length(x))))
split_data

```
###########
#4.c. Use unlist and matrix to transform the data to row-column format.
```{r}
data_matrix <- matrix(unlist(split_data), ncol = max_fields, byrow = TRUE)
data_matrix
```
###########
#4.d. From comment lines 2-4, extract the names of the fields. Set these as colnames for the matrix you just created.
```{r}
fieldNames <- gsub("^// Field [0-9]+: ", "", comments[2:4])
fieldNames

colnames(data_matrix) <- fieldNames
colnames(data_matrix)
```









